---
title: "Term Deposit Prediction & Marketing Analysis"
author: "Uche Kalu"
date: "4/9/2022"
output:
  pdf_document: default
  html_document: default
subtitle: Targeting Customers that are likely to opt into a Term Deposit service
---

**The goal of this analysis is to determine which features in the bank data relates to the bank's marketing campaign goal of having customers opt-into a TERM DEPOSIT (financial product)**

### **Executive Summary**

\- Duration since last contact and success of prior enrollments are correlated to success of Term Deposit opt-in

\- A strategy targeting customers that haven't been contacted in 319 days and have those with prior enrollments yields 32% positive outcomes (enrollments) vs 4.3% otherwise.

```{r include=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(readxl)
library(recipes)
library(tidyquant)
library(ggrepel)

```

### Analyzing and Investigating the Data

```{r save}
path   <- "bank_term_deposit_marketing_analysis.xlsx"
sheets <- excel_sheets(path)  # displays all d sheet names in d workbook
sheets
```

```{r include=FALSE}
sheets <- excel_sheets(path)  # displays all d sheet names in d workbook
sheets

sheets %>%
  map(~ read_excel(path  = path, sheet = .)) %>%  
  set_names(sheets)
```

**Our dataset parameters are in sheets 4 to 7 so we select only 4th to 7th sheets and combine all their data into one sheet.**

```{r results='hide'}
data_joined_tbl <- sheets[4:7] %>%   
  map(~ read_excel(path = path, sheet = .)) %>%
  reduce(left_join)   
```

```{r}
data_joined_tbl %>% glimpse()
```

### Analyze and Prepare the Data using recipe package

**Modeling and Machine learning often involves a response feature (Enrolled in TERM_DEPOSIT, yes/no) and many predictor features (AGE, JOB, MARITAL, etc). We are going to determine which predictors are related to the response variable. We do this using Binary Correlation Analysis**

```{r warning=FALSE, results='hide'}
recipe_obj <- recipe(~ ., data = data_joined_tbl) %>%  
  step_rm(ID) %>%                       
  step_discretize(all_numeric(), options = list(min_unique = 1)) %>%   
  step_dummy(all_nominal(), one_hot = TRUE, naming = partial(dummy_names, sep = "__")) %>% 
  prep()  

data_transformed_tbl <- data_joined_tbl %>%
  bake(recipe_obj, new_data = .) 
```

```{r}
data_transformed_tbl %>% glimpse()
```

### Correlation of various features to Term Deposit enrollment. 

**We perform a correlation analysis between the response variable(TERM_DEPOSIT_yes) and the rest of the features.**

```{r warning=FALSE, message=FALSE}
correlation_tbl <- data_transformed_tbl %>%  
  cor(y = data_transformed_tbl$TERM_DEPOSIT__yes) %>%  
  as_tibble(rownames = "feature") %>%  
  rename(TERM_DEPOSIT__yes = V1) %>%  
  separate(feature, into = c("feature", "bin"), sep = "__") %>%  
  filter(!is.na(TERM_DEPOSIT__yes)) %>%    # removes all NA values
  filter(!str_detect(feature, "TERM_DEP")) %>%  
  arrange(abs(TERM_DEPOSIT__yes) %>% desc()) %>%  
  mutate(feature = as_factor(feature) %>% fct_rev())  

correlation_tbl

```

### Visualize the Correlations Funnel using Scatterplot 

**Examining the result chart, we deduce that Duration since last contact and success of prior enrollments are correlated to success of Term Deposit opt-in.**

```{r warning=FALSE}
correlation_tbl %>%
  ggplot(aes(TERM_DEPOSIT__yes, y = feature, text = bin)) + 
  geom_vline(xintercept = 0, linetype = 2, color = "red") + 
  geom_point(color = "#2c3e50") +   # creates datapoints in d charts
  geom_text_repel(aes(label = bin), size = 3, color = "#2c3e50") +  
  expand_limits(x = c(-0.4, 0.4)) +  # increasing d range of values in x-axis
  theme_tq() +
  labs(title = "Bank Marketing Analysis",
       subtitle = "Correlations to Enrollment in Term Deposit",
       y = "", x = "Correlation to Term Deposit")

```

### Lets Interpret the Correlations 

**We have bins as values for Duration, Balance, Campaign, Day and Age in the chart and we need to know what values are in them bins?**

```{r}
recipe_obj %>% tidy()   

bins_tbl <- recipe_obj %>% tidy(2)
bins_tbl               
View(bins_tbl)

bins_tbl %>% filter(terms == "DURATION")  


```

### STRATEGY

**A strategy targeting customers that haven't been contacted in 319 days and have those with prior enrollments yields 32% positive outcomes (enrollments) vs 4.3% otherwise.**

**We focus on value - BIN4 in DURATION column and value - success in POUTCOME column since they are the top columns that affects Term Deposit enrollment and their stated values are the only values at far RHS**

```{r results='hide'}
strategy_tbl <- data_joined_tbl %>%  
  select(DURATION, POUTCOME, TERM_DEPOSIT) %>%  
  mutate(POTENTIAL = case_when(DURATION > 319 ~ "High Potential",
                               POUTCOME == "success" ~ "High Potential",  
                               TRUE ~ "Normal")) %>%    
  group_by(POTENTIAL) %>%  # displays unique values in d stated column
  count(TERM_DEPOSIT) %>%  
  mutate(prop = n / sum(n)) %>%  
  ungroup() %>%
  mutate(label_text = str_glue("n: {n}
                                 prop: {scales::percent(prop)}")) 
```

```{r}
strategy_tbl
```

### REPORT RESULTS using a stacked Bar chart

```{r}
strategy_tbl %>%
  ggplot(aes(POTENTIAL, prop, fill = TERM_DEPOSIT)) + 
  geom_col() +
  geom_label(aes(label = label_text), fill = "white", color = "#2c3e50") +  
  scale_fill_tq() +                      # changes d color of d stacked bars 
  scale_y_continuous(labels = scales::percent_format()) +  
  theme_tq() +
  labs(title = "Bank Marketing Strategy",
       subtitle = str_glue("Targeting customers that haven't been contacted in 319 days or those with prior enrollments yields 32% vs 4.3%"))
```
